Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-09-24 forecast/paper/supplement-text.tex
add statistical significance section to forecast paper supplement
- Some supplemental figures need some cleanup.
- Check on the naming of "all_signals_wide" in the Data Store.

>>>>

\section{Statistical significance}

In the Introduction of the main paper, we give some reasons that we avoid making
formal statements about statistical significance, preferring instead to examine
the stability of our results in different contexts. There are strong reasons to
avoid model-based significance tests because the necessary assumptions about
stationarity, independence, and the  model being true (or at least approximately
true) are certainly violated. With those caveats in mind, we undertake two
relatively assumption-lean investigations in this section. The first is a
sign-test for whether the difference between the AR model’s relative WIS  and
each other model’s relative WIS is centered at zero. By ``relative WIS'' we mean
scaled by the strawman as displayed in Figure 3. To mitigate the dependence
across time (which intuitively seems to matter more than that across space), we
computed these tests in a stratified way, where for each forecast date we run a
sign test on the scaled errors between two models over all 440 counties. The
results are plotted as histograms in Figure \ref{fig:sign-test}. In this case,
we use the total relative WIS over all aheads, but the histograms are largely
similar for individual target horizons. If there were no difference, we would
expect to see a uniform distribution. However, for all indicator-assisted
models, we see many more small $P$-values than would be expected if the null
hypothesis (that the AR model is better) were true.

Another relatively assumption-lean method of testing for differences in forecast
accuracy is the Diebold-Mariano test \cite{Diebold:2002,Diebold:2015,harvey1997testing}.
Essentially, the differences between forecast losses are assumed to have a
constant mean and a covariance that depends on time. Under these conditions, the
asymptotic distribution for the standardized mean of the differences is limiting
normal provided that a heteroskedasticity and autocorrelation robust estimate of
the variance is used. Using the loss as weighted interval score across all HRRs
and horizons (7 to 21 days ahead), we perform the DM test using both the mean
relative to the strawman (as reported in the manuscript) and the geometric mean
relative to the strawman as described above. The first two rows of Table
\ref{tab:dm-test} displays 
$P$-values for the test that the indicator-assisted model is no-better than the
AR model. Only the CHNG-CLI model's $P$-value exceeds conventional statistical
significance thresholds.

### 2021-09-24 forecast/paper/supplement.Rmd
add statistical significance section to forecast paper supplement
- Some supplemental figures need some cleanup.
- Check on the naming of "all_signals_wide" in the Data Store.

---
bibliography: [../../common/covidcast.bib, pnas-materials/pnas-sample.bib]
output: 
  pdf_document:
    keep_tex: true
    template: pnas-suppl-template.tex
params:
  fd_casefloor: 30
  flag_jumps: 15
---

@@

```{r hot-honest-v-finalized, fig.cap="Relative AUC with vintage compared to finalized data. Using finalized data leads to overly optimistic hotspot performance."}
left_join(
  hotspots_honest %>%
    group_by(forecaster, ahead) %>%
    summarise(auc_honest = auc(value, actual)),
  hotspots_finalized %>%
    group_by(forecaster, ahead) %>%
    summarise(auc_finalized = auc(value, actual))) %>%
  ggplot(aes(ahead, auc_honest / auc_finalized, color = forecaster)) + 
  geom_line() + 
  geom_point() +
  theme_bw() + 
  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow=1)) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  geom_hline(yintercept = 1) +
  ylab("AUC (vintage / finalized)")
```

@@

>>>>

---
bibliography: [../../common/covidcast.bib]
output: 
  pdf_document:
    keep_tex: true
    template: pnas-suppl-template.tex
params:
  fd_casefloor: 30
  flag_jumps: 15
---

@@

```{r hot-honest-v-finalized, fig.cap="Relative AUC with vintage compared to finalized data. Using finalized data leads to overly optimistic hotspot performance."}
left_join(
  hotspots_honest %>%
    filter(period != "jm") %>%
    group_by(forecaster, ahead) %>%
    summarise(auc_honest = auc(value, actual)),
  hotspots_finalized %>%
    filter(period != "jm") %>%
    group_by(forecaster, ahead) %>%
    summarise(auc_finalized = auc(value, actual))) %>%
  ggplot(aes(ahead, auc_honest / auc_finalized, color = forecaster)) + 
  geom_line() + 
  geom_point() +
  theme_bw() + 
  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow=1)) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  geom_hline(yintercept = 1) +
  ylab("AUC (vintage / finalized)")
```

@@

<!-- Statistical significance -->

```{r sign-test, fig.cap="P-values for a sign test that the WIS of the AR forecaster is smaller than that of the indicator-assisted model. Each P-value corresponds a particular forecast date."}
fcast_colors2 <- fcast_colors[names(fcast_colors) != "AR"]

st <- fcasts_honest %>%
  filter(period != "jm") %>%
  mutate(wis = wis / strawman_wis) %>%
  select(forecaster, geo_value, ahead, forecast_date, wis) %>%
  pivot_wider(names_from = forecaster, values_from = wis) %>%
  mutate(across(AR:`Google-AA`, ~ AR - .x)) %>%
  select(-AR) %>%
  pivot_longer(`CHNG-CLI`:`Google-AA`, names_to = "forecaster", values_to = "dif") %>%
  group_by(forecaster, forecast_date, geo_value) %>%
  summarise(dif = sum(dif)) %>%
  group_by(forecaster, forecast_date) %>%
  summarise(p = binom.test(
    x = sum(dif > 0),
    n = n(),
    alternative = "greater")$p.val)

ggplot(st, aes(p)) +
  geom_histogram(
    aes(y = after_stat(density), color = forecaster, fill = forecaster), 
    alpha = 0.4,
    bins = 40) +
  scale_color_manual(values = fcast_colors2, guide = guide_legend(nrow = 1)) +
  scale_fill_manual(values = fcast_colors2, guide = guide_legend(nrow = 1)) +
  facet_wrap(~forecaster) +
  ylab("Frequency") +
  xlab("P-value") +
  scale_y_continuous(
    labels = scales::percent_format(scale = 1, accuracy = 1)) +
  theme_bw() +
  theme(legend.position = "bottom")

```{r dm-test}
errs <- fcasts_honest %>% 
  filter(period != "jm") %>%
  group_by(forecaster, forecast_date) %>%
  summarise(
    `Mean Relative WIS` = mean(wis) / mean(strawman_wis),
    # mwis = mean(wis), 
    `Geometric Mean Relative WIS` = GeoMean(wis / strawman_wis) #,
    # gmwis = GeoMean(wis)
    ) %>%
  ungroup()
ar <- errs %>% 
  filter(forecaster == "AR") %>% 
  select(-forecaster) %>%
  pivot_longer(-forecast_date, names_to = "metric", values_to = "AR")
notar <- errs %>%
  filter(forecaster != "AR") %>% 
  pivot_longer(-c("forecaster", "forecast_date"), names_to = "metric", values_to = "err")

all_errs <- left_join(notar, ar)
dms <- all_errs %>% 
  group_by(forecaster, metric) %>% 
  summarise(dm = forecast::dm.test(AR, err, "greater", h = 7 , power = 1)$p.value)

kableExtra::kable(
  bind_rows(dms) %>% 
    pivot_wider(names_from = forecaster, values_from = dm), 
  digits = 3, booktabs = TRUE, caption = "$P$-values for a Diebold-Mariano test for differences in forecast error. The test is for the null hypothesis of equal performance against the alternative that the indicator assisted model is better.")
```

### 2021-09-24 forecast/paper/supplement.tex
add statistical significance section to forecast paper supplement
- Some supplemental figures need some cleanup.
- Check on the naming of "all_signals_wide" in the Data Store.

\author{Daniel J. McDonald, Jacob Bien, Alden Green, Addison Hu, Nat DeFries,
  Sangwon Hyun, Natalia L. Oliveira, James Sharpnack, Jingjing Tang, Robert
  Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J. Tibshirani} 
\correspondingauthor{Daniel J. McDonald.\\E-mail: daniel@stat.ubc.ca}

@@

@@

%\bibliography{../../common/covidcast.bib,pnas-materials/pnas-sample.bib}

>>>>

\author{Daniel J. McDonald, Jacob Bien, Alden Green, Addison J. Hu, Nat DeFries,
  Sangwon Hyun, Natalia L. Oliveira, James Sharpnack, Jingjing Tang, Robert
  Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J. Tibshirani} 
\correspondingauthor{Daniel J. McDonald.\\E-mail: daniel@stat.ubc.ca}

@@

{\centering \includegraphics[width=\textwidth]{fig/sign-test-1}

\caption{P-values for a sign test that the WIS of the AR forecaster is smaller than that of the indicator-assisted model. Each P-value corresponds a particular forecast date.}\label{fig:sign-test}
\end{figure}

\begin{table}

\caption{\label{tab:dm-test}$P$-values for a Diebold-Mariano test for differences in forecast error. The test is for the null hypothesis of equal performance against the alternative that the indicator assisted model is better.}
\centering
\begin{tabular}[t]{lrrrrr}
\toprule
metric & CHNG-CLI & CHNG-COVID & CTIS-CLIIC & DV-CLI & Google-AA\\
\midrule
Geometric Mean Relative WIS & 0.141 & 0.076 & 0.000 & 0.059 & 0.008\\
Mean Relative WIS & 0.186 & 0.137 & 0.002 & 0.074 & 0.064\\
\bottomrule
\end{tabular}
\end{table}

@@

\bibliography{../../common/covidcast.bib}
